---
title: "SQL City Murder Mystery"
author: "Van Stewart"
format: html
---

For this lab, you will be joining and filtering related data sets to solve a murder mystery!

# Part 1: GitHub Workflow


At the top of the document (in the YAML) there is an `author` line that says 
`"Your name here!"`. Change this to be your name and save your file either by
clicking on the blue floppy disk or with a shortcut (command / control + s). 

Be sure to [commit the files to your repo](https://happygitwithr.com/existing-github-first#stage-and-commit). 

Let's get started!


# Part 2: Some Words of Advice

-   Set chunk options carefully.

-   Make sure you don't print out more output than you need.

-   Make sure you don't assign more objects than necessary---avoid "object junk"
in your environment.

-   Make your code readable and nicely formatted.

-   Think through your desired result **before** writing any code.


# Part 3: Finding the Killer

Northwestern University's Knight Lab wanted to help sharpen users' database
skills, so they created a murder mystery. Can you solve this crime in SQL City??

The relational data you will be working with contains tables with different
pieces of information pertinent to the crime - people, social media check-ins,
driver's licenses, crime scene reports, police interviews, and more!


## Access the Data

This code chunk will read in **all** of the tables of data for you. Don't modify or
remove this! Take some time to look at each file type so that 

```{r}
#| message: false
#| warning: false
#| label: read-in-data

library(tidyverse)

# If purrr is not detected, install the package
if (!"purrr" %in% installed.packages()) install.packages("purrr")

source("https://raw.githubusercontent.com/jcanner/stat_210_2025_website/main/labs/instructions/lab-7-setup.R")
```

## Solve the Crime

### Crime Scene Report

Detective Wickham reaches out to you...

> A crime has taken place and I need your help! There was a murder in SQL City sometime on January 15, 2018. Could you retrieve the crime scene report from the police department’s database and follow the clues to find the person responsible for the murder?!


**Step 1: Find the police report in `crime_scene_report`. Then used the information in the report to move on to the next data source based on the information learned.**

```{r}
#| label: crime-scene-report

## Filter to find the murder that happened in SQL City on Jan 15, 2018

look_ <- crime_scene_report |>
  filter(date == "20180115", type == "murder", city == "SQL City") 

view(look_)
```




**Next Steps: Follow the evidence to the person responsible for the murder, building a report as you go.**
There are accomplices, some knowingly and some unknowingly, but there is only one mastermind.


Solve the murder mystery, showing **all of your work in this document**. Your
document and code must be well organized, easy to follow, and reproducible.

+ Use headers and written descriptions to indicate what you are doing.
+ Use `dplyr` verbs and `join` functions rather than just looking through the
tables manually. Functions from `stringr` and `lubridate` will also be useful.
+ Use good code formatting practices.
+ Comment your code.
+ Cite any external sources you use to solve the mystery.



```{r}
#| label: witnessFinding

## Looking for all people named annabel who live on franklin
witness1 <- person |>
  filter(str_detect(name, "Annabel") &  address_street_name == "Franklin Ave")


## last person who lives on franklin
witness2 <- person |>
  filter( address_street_name == "Franklin Ave")|>
  arrange(desc(address_number))|>
  slice(1)
                    
                    
          

```

```{r}
#| label: interviews

## Look up interviews from those two witnesses
interview |>
  filter(person_id %in% c(witness1$id, witness2$id))


```

```{r}
#| label: gymLook!

## see who was there
possiblekillers <- get_fit_now_check_in|>
  filter(check_in_date == "20180109")

possiblekillers2 <- get_fit_now_member|>
  filter( id %in% possiblekillers$membership_id)

```


```{r}
#|label: moreInterviews


## interview people
possiblekillers3 <- person |>
  filter(name %in% possiblekillers2$name)

interview|>
  filter( person_id %in% possiblekillers3$id)

## Someone was hired by a women with a lot of money...

```
```{r}
#|label: someone who drives a model s, and is either 5.5 or 5.7, and has gone to the play

## Filter drivers with red hair, Tesla Model S, and height 65–67 inches
red_tesla <- drivers_license |>
  filter(car_make == 'Tesla',
         car_model == 'Model S',
         height >= 65,
         height <= 67,
         hair_color == 'red')

## Join with person table to get individuals matching the above driver profile
red_tesla_people <- person |>
  inner_join(red_tesla, by = c("license_id" = "id"))

## Filter concert check-ins for SQL Symphony Concert during December 2017
concert_checkins <- facebook_event_checkin |>
  filter(event_name == "SQL Symphony Concert",
         date >= "20171201", date <= "20171231")

## Count how many times each person attended and filter for those who attended 3 times
concert_counts <- concert_checkins |>
  count(person_id) |>
  filter(n == 3)

## Find the person(s) who match both the Tesla profile and attended 3 times
final_suspect <- red_tesla_people |>
  filter(id %in% concert_counts$person_id)

## Display the final suspect's information
final_suspect




```


::: callout-note
## And the final suspect is...

*Miranda Priestly*

:::
